# 1. Data loading and cleaning

## Context

Streaming platforms have changed how we consume entertainment, and **Twitch** is the leader in live streaming, especially for video games. A close friend of mine became part of that phenomenon: what began as a hobby, streaming games and building a small community, has slowly grown. He is not famous yet, but he dreams of making a living from it.

I have always been curious about how Twitch works and, seeing his daily effort, I decided to help. The goal of this project is to analyse platform data to find which strategies, behaviours or patterns help a streamer stand out and grow faster. The key question is:

> **How can a streamer, especially in the early years, identify and prioritise the strategies that maximise growth on Twitch?**

The analysis is split into three notebooks:

1. Data loading and cleaning (this notebook)
2. Univariate exploratory analysis
3. Bivariate exploratory analysis

## Data loading and cleaning

Required packages: `readr`, `tidyverse`, `naniar`, `knitr`, `RColorBrewer`, `viridis`, `corrplot`.

```{r libraries, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(naniar)
library(knitr)
library(RColorBrewer)
library(viridis)
library(corrplot)
```

The dataset is the Kaggle dataset [Top 1000 Twitch Streamers Data (May 2024)](https://www.kaggle.com/datasets/hibrahimag1/top-1000-twitch-streamers-data-may-2024). Its variables are:

```{r data-dictionary, echo=FALSE}
data_dictionary <- data.frame(
  Variable = c("RANK", "NAME", "LANGUAGE", "TYPE", "MOST_STREAMED_GAME",
               "2ND_MOST_STREAMED_GAME", "AVERAGE_STREAM_DURATION",
               "FOLLOWERS_GAINED_PER_STREAM", "AVG_VIEWERS_PER_STREAM",
               "AVG_GAMES_PER_STREAM", "TOTAL_TIME_STREAMED", "TOTAL_FOLLOWERS",
               "TOTAL_VIEWS", "TOTAL_GAMES_STREAMED", "ACTIVE_DAYS_PER_WEEK",
               "MOST_ACTIVE_DAY", "DAY_WITH_MOST_FOLLOWERS_GAINED"),
  Description = c("Streamer ranking", "Streamer nickname",
                  "Stream language", "Stream type",
                  "Most streamed game", "Second most streamed game",
                  "Average stream duration (hours)", "Followers gained per stream",
                  "Average viewers per stream", "Average number of games per stream",
                  "Total time streamed (hours)", "Total followers",
                  "Total views", "Total number of games streamed",
                  "Active days per week", "Most active day",
                  "Day with most followers gained"))
kable(data_dictionary)
```

I will focus on the metrics that help answer the initial question: **language, most streamed games, most active days, average stream duration and followers gained per stream**. The data represents the top 1,000 streamers on the platform, so it gives a view of the general behaviour and trends that I can then use to draw my own conclusions.

```{r load-data, message=FALSE, warning=FALSE}
data_twitch <- read_csv("dataset/datasetV2.csv")
```

```{r head}
head(data_twitch, 20)
```

```{r missing-before}
miss_var_summary(data_twitch)
```

The dataset has **`r format(nrow(data_twitch), big.mark = ",")` rows and `r ncol(data_twitch)` variables**, and the column types are correctly defined.

Only one variable has missing values (`n_miss`): **2ND_MOST_STREAMED_GAME**. That makes sense: some streamers only ever play one game, so there is no second game. To make the column names easier to work with, I rename `2ND_MOST_STREAMED_GAME` (names starting with a digit are awkward in R), convert all names to lower case, and replace the missing values with the label **No second game**.

```{r rename}
data_twitch <- rename(data_twitch, SECOND_MOST_STREAMED_GAME = `2ND_MOST_STREAMED_GAME`)
names(data_twitch) <- str_to_lower(names(data_twitch))

data_twitch <- data_twitch %>%
  mutate(second_most_streamed_game = replace_na(second_most_streamed_game, "No second game"))
```

Check that the change was applied:

```{r check-no-game}
data_twitch %>%
  filter(second_most_streamed_game == "No second game") %>%
  summarise(streamers_without_second_game = n())
```

```{r missing-after}
miss_var_summary(data_twitch)
```

There are no missing values left, so the exploratory analysis can start.
